---
title: "Fitness Activity Prediction Using Random Forest"
date: "Wednesday, July 23, 2014"
output: html_document
---
This document briefly describes machine method used to predict fitness activity
class based on data gathered through Fitbit fitness montoring device. In the 
following, Method section describes preprocessing, method and prediction process
and Summary section finally summarizes the entire activity.

### Method
Original Fitbit training data (given in pml-training.csv) consisted of 19622 observations and 160 variables. Some variables had less than 5% observations. Therefore, we
decided to drop all those variable which had more than 19000 missing values. Additionally, we do want to use only numerical variables to train our method. That is why, we
also removed all character variables from training data. Following R code performed
these two tasks;
```{r echo=FALSE, warning=FALSE}
require(caret, warn.conflicts = FALSE, quietly=TRUE)
library(ggbiplot, warn.conflicts = FALSE, quietly=TRUE)
library(randomForest, warn.conflicts = FALSE, quietly=TRUE)
setwd("C:/Dropbox/datasci-jhi/machinelearning/course_project/")
data <- read.csv("pml-training.csv", stringsAsFactor=F)
data$classe <- as.factor(data$classe)

excl = vector()
for (i in 1:159) {
  if ((class(data[, i]) == "character") | (sum(is.na(data[, i])) > 19000)) {
    excl = append(excl, i)
  }
  if (class(data[, i]) == "integer"){
    data[, i] = as.numeric(data[, i])
  }
}
data = data[-excl]; ncol = ncol(data)
```
Afterward, we set aside 25% data from training set for validation purpose. The
code is as follows;
```{r}
inTrain <- createDataPartition(y=data$classe, p=0.75, list=F)
trainData <- data[inTrain, ]
valData <- data[-inTrain, ]
```

To construct our model based on random forest algorith, we used 20 iterations and 1000
trees as given below;

```{r cache=TRUE}
rf <-randomForest(classe~.,data=trainData, mtry=2, ntree=1000, 
                  keep.forest=TRUE, importance=TRUE)
```
Then we validated the model with 25% set aside from training set. The validation confirmed 99% accuracy of resultant model.

```{r}
pr = predict(rf, testData)
t = table (obs=testData$classe, predict=pr)
round(prop.table(t, 1), 4)
```
The smaple error was significantly lower than 0.01.

### Summary
Our random forest based machine learning algorithm is high accurate in prediction
of Fitness Activity. We were able to predict activity of 20 test cases with 100%
accuracy.

### Appendix 
Supported R code is given in pml-model-search.r and pml-final-prediction.r

